---
title: "wuha-srr-analysis-dataset.csv/.rds codebook"
subtitle: "WUHA 0-36m seroreversion study longitudinal public dataset"
date: "updated: `r Sys.time()`"
authors: "Ben Arnold (ben.arnold@ucsf.edu)"
output: 
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Here, you will find the codebook for the file `wuha-srr-analysis-dataset.csv/.rds`. This includes the longitudinal data for the WUHA trial during rounds 0-36. The dataset was created by `1-wuha-srr-data-processing.R`. There are two formats for the dataset, .csv and .rds butthe contents are the same. 

The Rmd script that creates the codebook is `2-make-codebook-wuha-srr-analysis-dataset.Rmd` which is 
called from the shell script `0-wuha-srr-RUNALL.sh`.

The code below  based on Ruben Arslan's article on the "codebook" package: https://journals.sagepub.com/doi/full/10.1177/2515245919838783.

```{r prepare_codebook, message = FALSE}
library(here)
source(here("R/0-wuha-srr-Config.R"))
library(codebook)

# WUHA SRR analysis dataset
wuha_df <- read_rds(file = here("data/2-final","wuha-srr-analysis-dataset.rds"))

```

```{r prepare_dataset_indiv, message = FALSE}

# "omit the following lines, if your missing values are already properly labelled" (R. Arslan)
wuha_df <- detect_missing(wuha_df,
    only_labelled = FALSE, # "only labelled values are autodetected as
                                   # missing
    negative_values_are_missing = FALSE, # negative values are missing values
    ninety_nine_problems = TRUE,   # 99/999 are missing values, if they
                                   # are more than 5 MAD from the median" (R. Arslan)
    )

# labelling variables
var_label(wuha_df) <- list(
                                       cluster_id_public = "Public cluster ID", 
                                       household_id_public = "Public household ID", 
                                       individual_id_public = "Public individual ID",
                                       survey = "Month of follow-up",
                                       arm = "Treatment group (randomized, cluster-level)",
                                       age_months = "Age in months", 
                                       age_years = "Age in years",
                                       sex = "Child sex",
                                       age_group = "Child age group in years (category)",
                                       population = "WUHA sampling population",
                                        pcr_individual = "Individual childPCR test result", 
                                       tf = "Presence of TF", ti = "Presence of TI", 
                                       pgp3_mfi = "Pgp3 Luminex median flourescence intensity minus background (MFI-bg)",
                                       pgp3_pos = "Pgp3 seropositive",
                                       pgp3_cutoff = "MFI value used to determine seropositivity"
                                 )

# labelling factor levels
val_labels(wuha_df$tf) <- c("Present" = 1, "Absent" = 0)
val_labels(wuha_df$ti) <- c("Present" = 1, "Absent" = 0)
val_labels(wuha_df$pgp3_pos) <- c("Seropositive" = 1, "Seronegative" = 0)
val_labels(wuha_df$population) <- c("Random sample 1-5y" = "A", 
                                    "Random sample 6-9y" = "B", 
                                    "Random sample 10+y" = "C", 
                                    "Random sample 1-9y after baseline" = "DE", 
                                    "Longitudinal cohort not selected as poart of random sample" = "F")

# "If you are not using formr, the codebook package needs to guess which items
# form a scale. The following line finds item aggregates with names like this:
# scale = scale_1 + scale_2R + scale_3R
# identifying these aggregates allows the codebook function to
# automatically compute reliabilities.
# However, it will not reverse items automatically." (R. Arslan)
wuha_df <- detect_scales(wuha_df)

```

Create codebook

```{r metadata and codebook indiv}

metadata(wuha_df)$name <- "wuha-srr-analysis-dataset"
metadata(wuha_df)$description <- "This is a codebook for the wuha-srr-analysis-dataset.csv/.rds files."
codebook(wuha_df)
```

