---
title: "Seroconversion and seroreversion in the WUHA longitudinal cohort"
author: "Christine Tedijanto christine.tedijanto@ucsf.edu and Ben Arnold ben.arnold@ucsf.edu"
date: "Last updated: `r Sys.time()`"
output:
  html_document:
    highlight: default
    theme: default
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


## Setup
```{r setup}
#----------------------------
# source the project's
# configuration file
#----------------------------
library(here)
source(here("R/0-wuha-srr-Config.R"))
```

## Load dataset
```{r load data}
#----------------------------
# read in the analysis dataset
# this was created by
# 1-wuha-srr-data-processing.R
#----------------------------
df <- read_rds(here("data/2-final/wuha-srr-analysis-dataset.rds"))

#----------------------------
# rename public ID vars to
# usual names
#----------------------------
df <- df %>%
  rename(cluster_id = cluster_id_public,
         household_id = household_id_public,
         individual_id = individual_id_public)

#----------------------------
# prepare data frame to 
# look at changes over one year
#----------------------------
temp_curr_df <- df %>% mutate(next_survey = survey + 12)
temp_next_df <- df %>%
  dplyr::select(-c(arm)) %>% 
  rename_at(vars(-c(cluster_id, household_id, individual_id)), ~paste0("next_", .))

next_df <- temp_curr_df %>%
  left_join(temp_next_df, by = c("cluster_id", "household_id", "individual_id", "next_survey")) %>% 
  # add person-time
  mutate(py_time = case_when(
    is.na(pgp3_pos) | is.na(next_pgp3_pos) ~ 0,
    # assume that seroreversion / seroconversion occur halfway through the year
    (pgp3_pos == 0 & next_pgp3_pos == 1) ~ 0.5,
    (pgp3_pos == 1 & next_pgp3_pos == 0) ~ 0.5,
    TRUE ~ 1))

```


## Longitudinal cohort summary statistics

```{r summary_stats}
# number of serology measurements
n_sero <- next_df %>% filter(!is.na(pgp3_pos)) %>% nrow()
n_sero # 4327

# number of children
n_children <- next_df %>% filter(!is.na(pgp3_pos)) %>% pull(individual_id) %>% unique() %>% length() 
n_children # 1511

# number of one-year intervals
n_oneyear <- next_df %>% 
  filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) 

nrow(n_oneyear) # 2428
n_oneyear$individual_id %>% unique() %>% length() # 1221
(n_oneyear$individual_id %>% unique() %>% length())/n_children # 80.8%

## prevalence stats
get_overall_prev <- function(input_var, input_survey){
  temp <- next_df %>% 
    filter(survey == input_survey) %>% 
    group_by(cluster_id) %>% 
    summarise(n_pos = sum(get(input_var)==1, na.rm = TRUE),
              n_total = sum(!is.na(get(input_var))),
              .groups = "drop") %>% 
    mutate(cluster_prev = n_pos / n_total)
  
  median(temp$cluster_prev)
}

# sero
get_overall_prev(input_var = "pgp3_pos", input_survey = 0) # 0.322
get_overall_prev(input_var = "pgp3_pos", input_survey = 12) # 0.372
get_overall_prev(input_var = "pgp3_pos", input_survey = 24) # 0.417
get_overall_prev(input_var = "pgp3_pos", input_survey = 36) # 0.559

# pcr
get_overall_prev(input_var = "pcr_individual", input_survey = 0) # 0.061
get_overall_prev(input_var = "pcr_individual", input_survey = 12) # 0.207
get_overall_prev(input_var = "pcr_individual", input_survey = 24) # 0.263
get_overall_prev(input_var = "pcr_individual", input_survey = 36) # 0.302

# tf
get_overall_prev(input_var = "tf", input_survey = 0) # 0.650
get_overall_prev(input_var = "tf", input_survey = 12) # 0.479
get_overall_prev(input_var = "tf", input_survey = 24) # 0.569
get_overall_prev(input_var = "tf", input_survey = 36) # 0.469
```

## Estimating seroconversion

```{r scr_nonparam}
scr_n <- next_df %>%
  filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>%
  filter(pgp3_pos == 0, next_pgp3_pos == 1) %>%
  nrow()

scr_py <- next_df %>%
  filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>%
  filter(pgp3_pos == 0) %>%
  pull(py_time) %>%
  sum()

print(paste("seroconversions observed:", scr_n, "over", scr_py, "person-years"))

scr_nonparam <- scr_n / scr_py

print(paste0("nonparametric estimate: ", round(scr_nonparam*100, digits = 0), " per 100 person-years"))
```

```{r scr_reg}
# seroconversion regression model
scr_reg_fit <- glmer(next_pgp3_pos ~ 1 + (1|cluster_id),
                     family = poisson(link = "log"),
                     offset = log(py_time),
                     data = next_df %>%
                       filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>% 
                       filter(pgp3_pos == 0))

scr_reg_df <- data.frame(
  method = "regression",
  scr = exp(summary(scr_reg_fit)$coefficients["(Intercept)", "Estimate"]),
  scr_lb = exp(confint(scr_reg_fit, level = 0.95, method = "Wald")["(Intercept)", "2.5 %"]),
  scr_ub = exp(confint(scr_reg_fit, level = 0.95, method = "Wald")["(Intercept)", "97.5 %"])
  )

print(paste0("Poisson regression estimate: ",
             round(scr_reg_df$scr*100, digits = 1),
             " (",
             round(scr_reg_df$scr_lb*100, digits = 1),
             ",",
             round(scr_reg_df$scr_ub*100, digits = 1),
             ") per 100 person-years"))
```

## Estimating seroreversion

```{r srr_nonparam}
next_df %>% filter(pgp3_pos == 1, !is.na(next_pgp3_pos)) %>% nrow()
next_df %>% filter(pgp3_pos == 1, next_pgp3_pos == 1) %>% nrow()
(next_df %>% filter(pgp3_pos == 1, next_pgp3_pos == 1) %>% nrow())/(next_df %>% filter(pgp3_pos == 1, !is.na(next_pgp3_pos)) %>% nrow())

srr_n <- next_df %>%
  filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>%
  filter(pgp3_pos == 1, next_pgp3_pos == 0) %>%
  nrow()

srr_py <- next_df %>%
  filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>%
  filter(pgp3_pos == 1) %>%
  pull(py_time) %>%
  sum()

print(paste("seroreversions observed:", srr_n, "over", srr_py, "person-years"))

srr_nonparam <- srr_n / srr_py

print(paste0("nonparametric estimate: ", round(srr_nonparam*100, digits = 0), " per 100 person-years"))
```

```{r srr_reg}
# seroreversion regression model
srr_reg_fit <- glmer(next_pgp3_neg ~ 1 + (1|cluster_id),
                     family = poisson(link = "log"),
                     offset = log(py_time),
                     data = next_df %>%
                       filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>% 
                       filter(pgp3_pos == 1) %>%
                       # need to flip 1/0 so that seroreversions are the outcome of interest
                       mutate(next_pgp3_neg = ifelse(next_pgp3_pos==1,0,1)))

srr_reg_df <- data.frame(
  method = "regression",
  srr = exp(summary(srr_reg_fit)$coefficients["(Intercept)", "Estimate"]),
  srr_lb = exp(confint(srr_reg_fit, level = 0.95, method = "Wald")["(Intercept)", "2.5 %"]),
  srr_ub = exp(confint(srr_reg_fit, level = 0.95, method = "Wald")["(Intercept)", "97.5 %"])
  )

print(paste0("Poisson regression estimate: ",
             round(srr_reg_df$srr*100, digits = 1),
             " (",
             round(srr_reg_df$srr_lb*100, digits = 1),
             ",", 
             round(srr_reg_df$srr_ub*100, digits = 1),
             ") per 100 person-years"))
```

### Subgroup analyses

### By survey

```{r survey, fig.width = 8.5}
srr_survey_reg_fit <- glmer(next_pgp3_neg ~ 1 + (1|cluster_id) + as.factor(survey),
                            family = poisson(link = "log"),
                            offset = log(py_time),
                            data = next_df %>%
                              filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>% 
                              filter(pgp3_pos == 1) %>% 
                              mutate(next_pgp3_neg = ifelse(next_pgp3_pos==1,0,1)))

srr_survey_df <- confint(srr_survey_reg_fit, level = 0.95, method = "Wald") %>% 
  as.data.frame() %>%
  rownames_to_column(var = "variable") %>% 
  filter(variable != ".sig01") %>% 
  # rename variables for figure
  mutate(Month = case_when(
    variable == "(Intercept)" ~ "0 to 12 (reference)",
    variable == "as.factor(survey)12" ~ "12 to 24",
    variable == "as.factor(survey)24" ~ "24 to 36")) %>%
  # add exponentiated coefficients
  left_join(summary(srr_survey_reg_fit)$coefficients %>% as.data.frame() %>% rownames_to_column(var = "variable"),
            by = "variable") %>% 
  mutate_at(vars("Estimate", "2.5 %", "97.5 %"), ~exp(.)) %>% 
  # add reference group
  add_row(Month = "0 to 12", "2.5 %" = 1, "97.5 %" = 1, "Estimate" = 1)

srr_survey_df %>% filter(Month != "0 to 12") %>% dplyr::select(Month, Estimate, "2.5 %", "97.5 %")
```

### By age

```{r age, fig.width = 8.5}
next_df <- next_df %>% 
  mutate(age_group_no0 = case_when(
    #age_years %in% c(0) ~ "0y",
    age_group == "0-5y" ~ "1-5y",
    age_group == "6-9y" ~ "6-9y",
    age_group == "10+y" ~ "10+y"))

srr_age_reg_fit <- glmer(next_pgp3_neg ~ 1 + (1|cluster_id) + as.factor(age_group_no0),
                         family = poisson(link = "log"),
                         offset = log(py_time),
                         data = next_df %>%
                           filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos),
                                  age_group != "10+y") %>% 
                           filter(pgp3_pos == 1) %>% 
                           mutate(next_pgp3_neg = ifelse(next_pgp3_pos==1,0,1)))

srr_age_df <- confint(srr_age_reg_fit, level = 0.95, method = "Wald") %>% 
  as.data.frame() %>%
  rownames_to_column(var = "variable") %>% 
  filter(variable != ".sig01") %>% 
  # rename variables for figure
  mutate(age_group = case_when(
    #variable == "(Intercept)" ~ "0y (reference)",
    variable == "(Intercept)" ~ "1-5y (reference)",
    variable == "as.factor(age_group_no0)6-9y" ~ "6-9y")) %>%
  # add exponentiated coefficients
  left_join(summary(srr_age_reg_fit)$coefficients %>% as.data.frame() %>% rownames_to_column(var = "variable"),
            by = "variable") %>% 
  mutate_at(vars("Estimate", "2.5 %", "97.5 %"), ~exp(.)) %>% 
  # add reference group
  add_row(age_group = "1-5y", "2.5 %" = 1, "97.5 %" = 1, "Estimate" = 1)

srr_age_df %>% filter(age_group != "1-5y") %>% dplyr::select(age_group, Estimate, "2.5 %", "97.5 %")
```

### By starting MFI-bg

```{r mfi, fig.width = 8.5}
srr_quant_reg_fit <- glmer(next_pgp3_neg ~ 1 + (1|cluster_id) + mfi_log10,
                           family = poisson(link = "log"),
                           offset = log(py_time),
                           data = next_df %>%
                             filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>% 
                             filter(pgp3_pos == 1) %>% 
                             mutate(mfi = ifelse(pgp3_mfi<=0, 1, pgp3_mfi),
                                    mfi_log10 = log10(mfi),
                                    mfi_log10_cat = cut(mfi_log10, breaks = 5)) %>% 
                             mutate(next_pgp3_neg = ifelse(next_pgp3_pos==1,0,1)))

srr_quant_df <- data.frame(method = "regression",
                           srr = exp(summary(srr_quant_reg_fit)$coefficients["mfi_log10", "Estimate"]),
                           srr_lb = exp(confint(srr_quant_reg_fit, level = 0.95, method = "Wald")["mfi_log10", "2.5 %"]),
                           srr_ub = exp(confint(srr_quant_reg_fit, level = 0.95, method = "Wald")["mfi_log10", "97.5 %"]))

print(paste0("Poisson regression estimate: ", round(srr_quant_df$srr, digits = 2),
             " (", round(srr_quant_df$srr_lb, digits = 2), ",", round(srr_quant_df$srr_ub, digits = 2), ")"))
```

Re-estimate SRR, using categories of MFI on the log10 scale. Since the cutoff for seropositivity was 1113 in this study, use 4 categories of log10 3, 3.5, 4, >4
```{r mfi category, fig.width = 8.5}
#--------------------------
# create a categorical var
# for MFI levels
# >1000, >10,000
#--------------------------
next_df2 <- next_df %>%
  filter(!is.na(pgp3_pos), !is.na(next_pgp3_pos)) %>% 
  filter(pgp3_pos == 1) %>% 
  mutate(mfi = ifelse(pgp3_mfi<=0, 1, pgp3_mfi),
         mfi_log10 = log10(mfi),
         mfi_log10_cat = cut(mfi_log10, breaks = c(3,3.5,4,5)),
         mfi_log10_cat = relevel(mfi_log10_cat, ref="(3,3.5]")) %>% 
  mutate(next_pgp3_neg = ifelse(next_pgp3_pos==1,0,1))
#--------------------------
# cross tab of seroreversion
# by starting MFI level
#--------------------------
table(next_df2$mfi_log10_cat, next_df2$next_pgp3_neg)

#--------------------------
# model-based SRR estimate
#--------------------------
srr_mficat_reg_fit <- glmer(next_pgp3_neg ~ 1 + (1|cluster_id) + mfi_log10_cat,
                           family = poisson(link = "log"),
                           offset = log(py_time),
                           data = next_df2)

srr_mficat_df <- confint(srr_mficat_reg_fit, level = 0.95, method = "Wald") %>% 
  as.data.frame() %>%
  rownames_to_column(var = "variable") %>% 
  filter(variable != ".sig01") %>% 
  # rename variables for figure
  mutate(log10mficat = case_when(
    variable == "(Intercept)" ~ "3 to 3.5 (reference)",
    variable == "mfi_log10_cat(3.5,4]" ~ "3.5 to 4",
    variable == "mfi_log10_cat(4,5]" ~ ">4")) %>%
  # add exponentiated coefficients
  left_join(summary(srr_mficat_reg_fit)$coefficients %>% as.data.frame() %>% rownames_to_column(var = "variable"),
            by = "variable") %>% 
  mutate_at(vars("Estimate", "2.5 %", "97.5 %"), ~exp(.)) %>% 
  # add reference group
  add_row(log10mficat = "3 to 3.5", "2.5 %" = 1, "97.5 %" = 1, "Estimate" = 1) %>%
  mutate(log10mficat=factor(log10mficat,levels=c("3 to 3.5 (reference)","3 to 3.5","3.5 to 4",">4"))) %>%
  arrange(log10mficat)

srr_mficat_df %>% filter(log10mficat != "3 to 3.5 (reference)") %>% dplyr::select(log10mficat, Estimate, "2.5 %", "97.5 %")

```

## Estimating decay rate and half-life

Code based on Arnold et al. in eLife Supplementary Information File 8. 

Under an exponential decay model where IgG level at time $t$ is $N_t$ and the decay rate parameter is $\lambda$,  $N_t = N_0e^{(-\lambda t)}$, $\lambda = \frac{-\log(N_t/N_0)}{t}$, and the half life is $t_{1/2}=\log(2)/\lambda$.

```{r estimate decay overall}
# limit to children who are PCR negative
# and who did not have increasing mfi-bg
#
# lambda = -log(N(t)/N(o))*(1/t); t1/2 = log2/lambda
# or
# t_1/2 = t / log_{1/2}(N(t)/N(o))
decay_df <- next_df %>%
  filter(next_pcr_individual %in% c(0, 0.5),
         !is.na(pgp3_mfi), !is.na(next_pgp3_mfi),
         pgp3_pos == 1,
         next_pgp3_mfi<=pgp3_mfi) %>% 
  mutate(lambda = -log(next_pgp3_mfi/pgp3_mfi)/1,
         half_life = log(2)/lambda)

summary(decay_df$lambda)
summary(decay_df$half_life)

# histogram of decay estimates for those that are <=20 years
# this excludes 29 children for whom estimates >20
length(decay_df$half_life[decay_df$half_life > 20])
hist(decay_df$half_life[decay_df$half_life <= 20])
```

```{r estimate decay in groups}
decay_df_2 <- decay_df %>%
  mutate(fold_change = next_pgp3_mfi/pgp3_mfi) %>% 
  mutate(fold_label = case_when(
    fold_change < 0.25 ~ ">4 fold decrease",
    fold_change < 0.5 ~ ">2 fold decrease",
    TRUE ~ "<2 fold decrease"))

decay_df_2 %>% count(fold_label)

decay_df_2 %>% 
  group_by(fold_label) %>% 
  summarise(count = n(),
            mean_lambda = mean(lambda),
            sd_lambda = sd(lambda),
            median_lambda = median(lambda),
            mean_half_life = mean(half_life),
            sd_half_life = sd(half_life),
            median_half_life = median(half_life),
            .groups = "drop")

decay_df_2 %>% filter(fold_label == ">4 fold decrease") %>% pull(half_life) %>% summary()
```

```{r}
# are more rapid drops associated with younger age groups?
fold_age_model <- glm(fold_bin~age_years,
    family = binomial("logit"),
    data = decay_df_2 %>% 
      mutate(fold_bin = ifelse(fold_label == ">4 fold decrease", 1, 0)))

decay_df_2 %>% 
  ggplot(aes(x = age_years, fill = fold_label)) +
  geom_bar(stat = "count", position = "fill")
```

## Figures

### A. Counts and seroprevalence by month

```{r}
fig_counts_df <- next_df %>%
  count(survey, pgp3_pos) %>%
  mutate(pgp3_pos_label = case_when(
    pgp3_pos == 0 ~ "Negative",
    pgp3_pos == 1 ~ "Positive",
    TRUE ~ "Missing")) %>% 
  mutate(pgp3_pos_label = factor(pgp3_pos_label, levels = c("Missing", "Positive", "Negative")),
         survey = as.factor(survey))

fig_counts <- fig_counts_df %>% 
  ggplot(aes(x = survey, y = n, fill = pgp3_pos_label)) +
  geom_bar(stat = "identity", position = "stack",
           width = 0.6, color = NA) +
  geom_text(data = fig_counts_df %>%
              left_join(fig_counts_df %>%
                          group_by(survey) %>%
                          summarise(total = sum(n)),
                        by = "survey") %>%
              mutate(prop = n/total, prop_label = paste0(round(prop, digits=2)*100, "%")) %>%
              group_by(survey) %>% 
              # place text label in bar
              mutate(loc_y = cumsum(n) - 0.5*n) %>% 
              filter(pgp3_pos_label != "Missing"),
            aes(x = survey, y = loc_y, label = prop_label),
            color = "white", size = 2.5) +
  # based on Paul Tol's high contrast
  scale_fill_manual(values = c("Negative" = "#004488",
                               "Positive" = "#BB5566",
                               "Missing" = "#DDAA33")) +
  labs(x = "Study month", y = "Number of children", fill = "Pgp3 serostatus") +
  coord_cartesian(ylim = c(0,1500)) +
  theme_bw() +
  theme(legend.position = c(0.72, 0.86),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.key.height = unit(0.3, "cm"),
        plot.margin = margin(0.6, 0.2, 0.4, 0.5, "cm"))
```

### B. MFI-bg by PCR status

```{r mfi by pcr, warning = FALSE}
fig_mfi <- next_df %>%
  filter(pgp3_pos == 1, !is.na(next_pgp3_pos), !is.na(next_pcr_individual)) %>% # as of 10/08 update, remove missing PCR at x+12
  dplyr::select(individual_id, age_years, survey, pgp3_pos, next_pgp3_pos, pgp3_mfi, next_pgp3_mfi) %>% 
  # temp labels
  mutate(fold_change = next_pgp3_mfi/pgp3_mfi) %>% 
  mutate(fold_label = case_when(
    fold_change < 0.25 ~ ">4 fold decrease",
    fold_change < 0.5 ~ ">2 fold decrease",
    TRUE ~ "<2 fold decrease")) %>% 
  # end temp labels
  mutate(row_n = row_number()) %>% 
  pivot_longer(cols = c(pgp3_mfi, next_pgp3_mfi)) %>%
  left_join(next_df %>% dplyr::select(individual_id, survey, next_pcr_individual),
            by = c("individual_id", "survey")) %>%
  mutate(next_pcr_label = case_when(
    is.na(next_pcr_individual) ~ "PCR missing at x+12",
    next_pcr_individual == 0 | next_pcr_individual == 0.5  ~ "PCR negative/equivocal at x+12",
    next_pcr_individual == 1 ~ "PCR positive at x+12")) %>% 
  mutate(next_pcr_label = factor(next_pcr_label, levels = c("PCR positive at x+12", "PCR negative/equivocal at x+12", "PCR missing at x+12"))) %>% 
  mutate(x_val = ifelse(name == "pgp3_mfi", 0, 1)) %>% 
  #filter(age_years>1) %>% 
  #ggplot(aes(x = x_val, y = value, group = row_n, color = next_pcr_label, lty = fold_label)) +
  ggplot(aes(x = x_val, y = value, group = row_n, color = next_pcr_label)) +
  geom_line(data = . %>% filter(next_pgp3_pos==1), color = "darkgrey", lwd = 0.3, alpha = 0.3) +
  geom_line(data = . %>% filter(next_pgp3_pos==0), alpha = 0.6, lwd = 0.5) +
  geom_hline(yintercept = 1113, lty = "dashed") +
  # keep as boxplot instead of violin bc several outliers cause violin to be elongated
  geom_boxplot(aes(x = x_val+c(-0.07,0.07), y = value, group = x_val),
               inherit.aes = FALSE,
               fill = NA, width = 0.09, outlier.shape = NA,
               lwd = 0.4) +
  scale_x_continuous(breaks = c(0,1), labels = c("x", "x+12")) +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x)(c(1, 1e5)),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_color_manual(values = c("PCR negative/equivocal at x+12" = "#004488",
                                "PCR positive at x+12" = "#BB5566",
                                "PCR missing at x+12" = "#DDAA33")) +
  #scale_linetype_manual(values = c(">4 fold decrease" = "solid", ">2 fold decrease" = "dashed", "<2 fold decrease" = "dotted")) +
  coord_cartesian(xlim = c(-0.1,1.1)) +
  facet_grid(cols = vars(next_pcr_label)) +
  labs(y = "Pgp3 IgG level (MFI-bg)", x = "Study month") +
  theme_bw() +
  theme(#panel.grid = element_blank(),
        #axis.text.x = element_text(face = "italic"),
        legend.position = "none",
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size = 10),
        plot.margin = margin(0.6, 0.2, 0.4, 0.6, "cm"))
```

### C. Seroreversion by survey

```{r}
fig_surveys <- srr_survey_df %>% 
  filter(Month != "0 to 12 (reference)") %>% 
  ggplot() +
  geom_hline(yintercept = 1, lty = "dashed") +
  geom_errorbar(aes(x = Month, ymin = get("2.5 %"), ymax = get("97.5 %")), width = 0) +
  geom_point(aes(x = Month, y = Estimate), pch = 21, size = 3, fill = "white") +
  scale_y_continuous(trans = "log", breaks = c(0.01, 0.1, 0.5, 1, 2)) +
  coord_cartesian(ylim = c(0.003, 2)) +
  annotate("text",x=1,y=0.6,label="ref") +
  labs(x = "Study month", y = "Seroreversion rate ratio") +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.margin = margin(0.6, 0.2, 0.2, 0.6, "cm"))
```

### D. Seroreversion by age group

```{r}
fig_ages <- srr_age_df %>% 
  filter(age_group != "1-5y (reference)") %>% 
  ggplot() +
  geom_hline(yintercept = 1, lty = "dashed") +
  geom_errorbar(aes(x = age_group, ymin = get("2.5 %"), ymax = get("97.5 %")), width = 0) +
  geom_point(aes(x = age_group, y = Estimate), pch = 21, size = 3, fill = "white") +
  scale_y_continuous(trans = "log", breaks = c(0.01, 0.1, 0.5, 1, 2)) +
  coord_cartesian(ylim = c(0.003, 2)) +
  annotate("text",x=1,y=0.6,label="ref") +
  labs(x = "Age group", y = "Seroreversion rate ratio") +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.margin = margin(0.6, 0.2, 0.2, 0.6, "cm"))
```

### E. Seroreversion by starting MFI-bg
```{r}
fig_mficat <- srr_mficat_df %>% 
  filter(log10mficat != "3 to 3.5 (reference)") %>% 
  ggplot() +
  geom_hline(yintercept = 1, lty = "dashed") +
  geom_errorbar(aes(x = log10mficat, ymin = get("2.5 %"), ymax = get("97.5 %")), width = 0) +
  geom_point(aes(x = log10mficat, y = Estimate), pch = 21, size = 3, fill = "white") +
  scale_y_continuous(trans = "log", breaks = c(0.01, 0.1, 0.5, 1, 2)) +
  coord_cartesian(ylim = c(0.003, 2)) +
  annotate("text",x=1,y=0.6,label="ref") +
  labs(x = "log10 IgG (MFI-bg)", y = "Seroreversion rate ratio") +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.margin = margin(0.6, 0.2, 0.2, 0.6, "cm"))
```

### Composite figure 

(original version not used)

```{r, fig.width = 8.5, fig.height = 6}
combined_top <- plot_grid(fig_counts, fig_mfi,
                          labels = "AUTO",
                          nrow = 1,
                          rel_widths = c(0.33, 0.66))
 
combined_bottom <- plot_grid(fig_surveys, fig_ages, fig_mficat,
                             nrow = 1, labels = c("C", "D","E"),
                             rel_widths = c(0.33,0.33,0.33))

combined_fig <- plot_grid(combined_top, combined_bottom,
                          ncol = 1, rel_heights = c(1, 0.8), labels = NULL)

combined_fig

ggsave(filename = here("output", "wuha_long_fig1_notused.png"),
       plot = combined_fig,
       height = 6, width = 8)
```

### Figure 1

```{r alt fig 1}
mfi_df <- next_df %>%
  filter(pgp3_pos == 1, !is.na(next_pgp3_pos), !is.na(next_pcr_individual)) %>% # as of 10/08 update, remove missing PCR at x+12
  dplyr::select(individual_id, age_years, survey, pgp3_pos, next_pgp3_pos, pgp3_mfi, next_pgp3_mfi) %>% 
  # temp labels
  mutate(fold_change = next_pgp3_mfi/pgp3_mfi) %>% 
  mutate(fold_label = case_when(
    fold_change < 0.25 ~ ">4 fold decrease",
    fold_change < 0.5 ~ ">2 fold decrease",
    TRUE ~ "<2 fold decrease")) %>% 
  pivot_longer(cols = c(pgp3_mfi, next_pgp3_mfi)) %>%
  mutate(interval_time = ifelse(name == "pgp3_mfi", 0, ifelse(name == "next_pgp3_mfi", 1, NA_real_))) %>% 
  left_join(next_df %>% dplyr::select(individual_id, survey, next_pcr_individual),
            by = c("individual_id", "survey"))

get_fig_mfi <- function(input_survey, input_pcr_pos){
  
  if(input_pcr_pos == TRUE){
    pcr_list <- c(1) # positive
  } else {
    pcr_list <- c(0.5, 0) # negative or equivocal
  }
  
  mfi_df %>% 
    filter(survey == input_survey, next_pcr_individual %in% pcr_list) %>% 
    ggplot(aes(x = interval_time, y = value, group = individual_id)) +
    geom_line(data = . %>% filter(next_pgp3_pos==1), color = "darkgrey", lwd = 0.3, alpha = 0.3) +
    geom_line(data = . %>% filter(next_pgp3_pos==0), alpha = 0.6, lwd = 0.5, color = "#BB5566") +
    geom_hline(yintercept = 1113, lty = "dashed") +
    # keep as boxplot instead of violin bc several outliers cause violin to be elongated
    geom_boxplot(aes(x = interval_time+c(-0.07,0.07), y = value, group = interval_time),
                 inherit.aes = FALSE,
                 fill = NA, width = 0.09, outlier.shape = NA,
                 lwd = 0.4) +
    scale_x_continuous(breaks = c(0, 1), labels = c(input_survey, input_survey+12)) +
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x)(c(1, 1e5)),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    coord_cartesian(xlim = c(0-0.1,1+0.1), ylim = c(1, 1e5)) +
    labs(y = "Pgp3 IgG level (MFI-bg)", x = "Study month") +
    theme_bw() +
    theme(legend.position = "none",
          strip.background = element_rect(fill="white"),
          strip.text = element_text(size = 9))#,
          #plot.margin = margin(0.6, 0.2, 0.4, 0.6, "cm"))
}

fig_mfi_pos <- plot_grid(get_fig_mfi(input_survey = 0, input_pcr_pos = TRUE),
                         get_fig_mfi(input_survey = 12, input_pcr_pos = TRUE),
                         get_fig_mfi(input_survey = 24, input_pcr_pos = TRUE),
                         labels = NULL,
                         nrow = 1,
                         rel_widths = c(0.33, 0.33, 0.33))
 
fig_mfi_neg_equi <- plot_grid(get_fig_mfi(input_survey = 0, input_pcr_pos = FALSE),
                              get_fig_mfi(input_survey = 12, input_pcr_pos = FALSE),
                              get_fig_mfi(input_survey = 24, input_pcr_pos = FALSE),
                              labels = NULL,
                              nrow = 1,
                              rel_widths = c(0.33, 0.33, 0.33))

alt_fig_1 <- plot_grid(fig_mfi_pos, fig_mfi_neg_equi,
                       ncol = 1, rel_heights = c(1, 1), labels = c("A", "B"))

alt_fig_1

ggsave(filename = here("output", "wuha_long_fig_1.png"),
       plot = alt_fig_1,
       height = 5, width = 8)
```

```{r alt fig 1 v2}
# try version of alt fig 1 that includes children missing PCR at x+12, and combines panels so that lines are colored by PCR status
# resulting figures look a bit cluttered - could consider just coloring those who serorevert, and color by PCR status
mfi_df <- next_df %>%
  filter(pgp3_pos == 1, !is.na(next_pgp3_pos)) %>% # as of 10/08 update, remove missing PCR at x+12
  dplyr::select(individual_id, age_years, survey, pgp3_pos, next_pgp3_pos, pgp3_mfi, next_pgp3_mfi) %>% 
  # temp labels
  mutate(fold_change = next_pgp3_mfi/pgp3_mfi) %>% 
  mutate(fold_label = case_when(
    fold_change < 0.25 ~ ">4 fold decrease",
    fold_change < 0.5 ~ ">2 fold decrease",
    TRUE ~ "<2 fold decrease")) %>% 
  pivot_longer(cols = c(pgp3_mfi, next_pgp3_mfi)) %>%
  mutate(interval_time = ifelse(name == "pgp3_mfi", 0, ifelse(name == "next_pgp3_mfi", 1, NA_real_))) %>% 
  left_join(next_df %>% dplyr::select(individual_id, survey, next_pcr_individual),
            by = c("individual_id", "survey"))

get_fig_mfi <- function(input_survey){
  
  mfi_df %>% 
    mutate(pcr_label = case_when(
      is.na(next_pcr_individual) ~ "Missing",
      next_pcr_individual == 1 ~ "Positive",
      next_pcr_individual %in% c(0, 0.5) ~ "Negative/equivocal"
    )) %>% 
    filter(survey == input_survey) %>% 
    ggplot(aes(x = interval_time, y = value, group = individual_id, color = pcr_label)) +
    geom_line(lwd = 0.3, alpha = 0.3) +
    geom_hline(yintercept = 1113, lty = "dashed") +
    # keep as boxplot instead of violin bc several outliers cause violin to be elongated
    geom_boxplot(aes(x = interval_time+c(-0.07,0.07), y = value, group = interval_time),
                 inherit.aes = FALSE,
                 fill = NA, width = 0.09, outlier.shape = NA,
                 lwd = 0.4) +
    scale_x_continuous(breaks = c(0, 1), labels = c(input_survey, input_survey+12)) +
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x)(c(1, 1e5)),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    scale_color_manual(values = c("Negative/equivocal" = "#004488",
                                  "Positive" = "#BB5566",
                                  "Missing" = "darkgrey")) + 
    coord_cartesian(xlim = c(0-0.1,1+0.1), ylim = c(1, 1e5)) +
    labs(y = "Pgp3 IgG level (MFI-bg)", x = "Study month") +
    theme_bw() +
    theme(legend.position = "none",
          strip.background = element_rect(fill="white"),
          strip.text = element_text(size = 9))#,
          #plot.margin = margin(0.6, 0.2, 0.4, 0.6, "cm"))
}

alt_fig_1_v2 <- plot_grid(get_fig_mfi(input_survey = 0),
                         get_fig_mfi(input_survey = 12),
                         get_fig_mfi(input_survey = 24),
                         labels = NULL,
                         nrow = 1,
                         rel_widths = c(0.33, 0.33, 0.33))

alt_fig_1_v2
```



### Figure 2

```{r alt fig 2}
alt_fig_2 <- plot_grid(fig_surveys, fig_ages, fig_mficat,
                       nrow = 1, labels = c("A", "B","C"),
                       rel_widths = c(0.33,0.33,0.33))

alt_fig_2

ggsave(filename = here("output", "wuha_long_fig_2.png"),
       plot = alt_fig_2,
       height = 2.5, width = 8)
```

### Supplementary Table 1

```{r table s1}
# number of children sampled at each survey time point
n_children_by_survey <- next_df %>% 
  filter(!is.na(pgp3_mfi)) %>% 
  group_by(survey) %>% 
  count(name = "n_children")

# age distribution by survey time point
age_by_survey <- next_df %>% 
  filter(!is.na(pgp3_mfi)) %>% 
  group_by(survey) %>% 
  summarise(
    age_median = median(age_months),
    age_q25 = quantile(age_months, p = 0.25),
    age_q75 = quantile(age_months, p = 0.75)
  )

# seroprevalence by survey time point
seroprev_by_survey <- next_df %>% 
  filter(!is.na(pgp3_mfi)) %>% 
  count(survey, pgp3_pos) %>%
  group_by(survey) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  mutate(seroprev = n/total) %>% 
  filter(pgp3_pos == 1) %>% 
  dplyr::select(-c(n, total, pgp3_pos))

# number of one-year intervals at each time point
n_oneyear_by_survey <- next_df %>% 
  filter(!is.na(pgp3_mfi), !is.na(next_pgp3_mfi)) %>% 
  group_by(next_survey) %>% 
  count() %>% 
  rename(survey = next_survey,
         n_oneyear = n)

table_s1 <- plyr::join_all(
  list(n_children_by_survey, age_by_survey, seroprev_by_survey, n_oneyear_by_survey),
  by = "survey",
  type = "left"
)

table_s1
```

## Session Info
```{r session info}
sessionInfo()
```